Data Preparation

### 2023-03-22 x_bra-check.qmd
still playing around with pre-processing

```{r}
library(tidyverse)
```

@@

```{r}
#sbgr_arp   <- c(rwys_sbgr$LON, rwys_sbgr$LAT)
sbgr_200NM <- trrrj::polygon_at_distance(c(8.570556, 50.03333), d = 205)
```

@@

>>>>

```{r}
library(tidyverse)
library(patchwork)
```

@@

```{r}
sbgr_arp   <- rwys_sbgr |> filter(ICAO == "SBGR", REF == "ARP") |> select(LAT, LON)
sbgr_200NM <- trrrj::polygon_at_distance(c(sbgr_arp$LON, sbgr_arp$LAT), d = 205) |> 
  sf::st_as_sf() |> 
  rename(geometry = x) |> 
  mutate(NAME = "SBGR-200")
```

```{r}
sbgr_base_plot <- ggplot() +
  geom_sf(data = sbgr_200NM) +
  geom_point(data = sbgr_arp, aes(x = LON, y = LAT))
sbgr_base_plot
```

@@

## spatial subsetting for SBGR

```{r}
spatial_subset <- function(.payload, .outer, ...){
  # check if payload is already a sf object
  if(!is(.payload,"sf")){
    .payload <- .payload |> cast_latlon_to_pts(.drop_coord = FALSE)
  }
  # spatial subsetting
  trimmed_data <- .payload[.outer,]
}

trjs_pts <- spatial_subset(bra |> filter(!is.na(LAT)), sbgr_200NM)

# trjs_pts |> sf::st_drop_geometry() |> 
#   arrow::write_parquet("./data/bra-sbgr-20220509.parquet")
```

```{r}
sbgr_base_plot + geom_sf(data = trjs_pts)
```

```{r}
picks <- trjs_pts |> filter(FLTID %in% c("TAM3201", "GLO1612","TAM3172", "DLH506"))

sbgr_base_plot + geom_sf(data = picks, aes(color = ALT_B))
```

```{r}
trjs_pts |> filter(FLTID == "DLH506") |> ggplot() + geom_point(aes(x = TIME, y = ALT_B))
```

## Subsetting for KATL

```{r}
katl_arp   <- c(-84.4278640,33.6366996)
katl_200NM <- trrrj::polygon_at_distance(katl_arp, d = 205) |> sf::st_as_sf() |> 
  rename(geometry = x) |> 
  mutate(NAME = "KATL-205")

katl_pts <- spatial_subset(bra |> filter(!is.na(LAT)), katl_200NM)
katl_pts |> sf::st_drop_geometry() |> arrow::write_parquet("./data/usa-katl-20220509.parquet")
```

```{}
# just a black ball = fully covered
ggplot() +
  geom_sf(data = katl_200NM) +
  geom_sf(data = katl_pts)
```

```{r}
check_flight <- katl_pts |> filter(ICAO24 %in% c("a3b4b5"))

p1 <- check_flight |> 
  ggplot() +
  geom_sf(data = katl_200NM) +
  geom_point(aes(x = LON, y = LAT, color = ALT_B))

p2 <- check_flight |> 
  ggplot() +
  geom_point(aes(x = TIME, y = ALT_B, color = ALT_B))

p1 + p2 + plot_layout(guides = 'collect')
```

### 2023-03-22 x_data-crunching.qmd
still playing around with pre-processing

# load list of chosen dumps
ds  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn() |> 
  mutate(ALT_B = coerce_meter_to_feet(ALT_B))
```

@@

rwy_centerline_ls <- cast_rwy_ctr_line_ls(rwys_egll)
rwy_box           <- airport_threshold_box(rwys_egll)
```

```{r}
apt_arrival_plot <- ggplot() +
  geom_sf(data = rq) +
  geom_sf(data = rwy_box) +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF)) 
apt_arrival_plot
```

apt_arrival_plot +
  geom_sf(data = baw236_landing_ls)

>>>>

# load list of chosen dumps
ds  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn() |> 
  mutate(across(.cols = c("ALT_B","ALT_G"), .fns = coerce_meter_to_feet))
```

## some trajectory cleaning

we study airborn part, i.e. from take-off to landing

* departing flights: velocity > 50 m/s
* landing flights cannot have altitude above 5000ft, velocity < 90

```{r}
ds  |> 
  filter(ICAO24 == "3c6498") |> 
  filter(velocity > 50) |> 
  Alt_Outlier_filter(.alt_var = ALT_B) |> 
  ggplot() + 
    geom_point(aes(x = TIME, y = ALT_B, color = velocity < 90 & ALT_B > 5000)) + 
    geom_point(aes(x = TIME, y = ALT_G), color = "green") +
  theme(legend.position = "top")
```

@@

rwy_centerline_ls <- cast_rwy_ctr_line_ls(rwys_egll)
rwy_box           <- airport_threshold_box(rwys_egll)
airport_box       <- airport_centerline_box(rwys_egll)
```

Check some of the functions and shapes

```{r}
apt_arrival_plot <- ggplot() +
  geom_sf(data = airport_box) +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF)) 
apt_arrival_plot
```

Address[which(LeastNEmployees == max(LeastNEmployees))]

```{r}
icao24_in_polygon <- function(.ds, .poly, .id = ICAO24){
  # determine bounding box for polygon
  this_box <- sf::st_bbox(.poly)
  # pre-filter with dplyr - cut bounding box
  icao24s <- .ds |> 
    dplyr::filter(
       LAT > this_box$ymin, LAT < this_box$ymax
      ,LON > this_box$xmin, LON < this_box$xmax, 
    )
  # get ids seen
  icao24s <- icao24s |> 
    dplyr::summarise(
      N = n(), N_ID = length(unique(FLTID))
      , MIN_TIME = min(TIME,na.rm = TRUE)
      , MIN_TIME_ALT = ALT_B[which(TIME == MIN_TIME)]
      , MAX_TIME = max(TIME, na.rm = TRUE)
      , MAX_TIME_ALT = ALT_B[which(TIME == MAX_TIME)]
      , MIN_ALT = min(ALT_B, na.rm = TRUE)
      , MED_ALT = median(ALT_B, na.rm = TRUE)
      , MAX_ALT = max(ALT_B, na.rm = TRUE)
      , .by = {{.id}})
}
```

```{r}
seen_at_airport <- icao24_in_polygon(ds, airport_box)
seen_at_airport
```

```{r}
source("~/RProjects/paper-2023-ICNS/R/trj_identify_legs.R")
source("~/RProjects/paper-2023-ICNS/R/coerce_unixepoch_to_datetime.R")
source("~/RProjects/paper-2023-ICNS/R/trj_fix_fltid.R")
source("~/RProjects/paper-2023-ICNS/R/add_distance_rowwise.R")
source("~/RProjects/paper-2023-ICNS/R/trj_fix_altitude.R")
source("~/RProjects/paper-2023-ICNS/R/heuristic_DAO.R")

# helper function to check missing FLTID or gaps
check_leg_fltid <- function(.trjs_with_legs, ...){
  .trjs_with_legs |> 
    dplyr::summarise(
        N = n(), N_ID = length(unique(FLTID))
      , IDs = paste(unique(FLTID), collapse = "-")
      , NAs = sum(is.na(FLTID))
      , .by = c("ICAO24","LEG")
      )
}
```

```{r}
rq <- ds |> 
  filter(ICAO24 %in% seen_at_airport$ICAO24) |> 
  coerce_unixepoch_to_datetime() |> 
  identify_trajectory_legs() |> 
  Alt_Outlier_filter(.alt_var = ALT_B) |> 
  fix_altitude_heuristic() |> 
  mutate(UID = paste(ICAO24, LEG, sep = "-"))
rq <- rq |> ungroup()

# check one more time for airport
seen_at_airport2 <- icao24_in_polygon(rq, airport_box, UID) |> 
  heuristic_DAO()
seen_at_airport2
```

```{r}
viz_check_DAO <- function(.trj, .apt_box = airport_box){
  plt <- .trj |> 
    ggplot() + 
      geom_sf(data = .apt_box) + 
      geom_point(data = .trj, aes(x = LON, y = LAT, color = TIME)) + 
      coord_sf(xlim = c(-1, 0), ylim = c(51.3, 51.6), expand = FALSE)
  return(plt)
}

picks <- seen_at_airport2$UID[1:9]

rq |> filter(UID %in% picks) |> viz_check_DAO() + facet_wrap(.~UID)
```
